Multiresolution Neural Field Estimation

### 2011-02-28 Multiresolution_framework/ltx/manuscript.tex
edits to paper

\section{Introduction}
\IEEEPARstart{T}{he} human brain has an inherent multiresolution architecture. Spatial scales of information processing and transfer range from individual proteins in synapses, to neurons, to networks of neurons. The multiresolution cortical architecture poses major modeling challenges to efficiently describe the brain's dynamics. This paper introduces a multiresolution data-driven framework for neural field modeling, the multiresolution approximate integrodifference equation (MRAIDE), to address this challenge. Neural field models typically describe the meso and macroscopic dynamics of the human brain, but are also descriptive of finer-scale neurodynamics.

The ability to create patient-specific neural models will contribute to our knowledge base of poorly understood diseases such as epilepsy, and furthermore, may also enable the development of new treatment strategies. This is particularly poignant with the advent of new devices that use therapeutic electrical stimulation. Presently, stimulation strategies for devices using corrective or modulating electrical stimulation operate in an open-loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists an enormous potential to improve the performance of these devices using systems theory. The design of a stimulation strategy using systems theory requires a model. Neural mass and neural field models are the ideal candidates for estimation and control algorithms. It is expected that parameters of the neural fields models will be patient specific and will therefore need to be inferred from data.

Creating patient specific neural models has been approached from a number of different standpoints. \dean{need to add proper citations here.} Valdes et al. (1999) used the neural mass model of Lopes Da Silva (1976) and Zetterberg et al. (1978) to fit EEG data. This approach was extended by David and Friston (2003) to coupled neural masses (using the equations of Jansen and Ritt (1995)) through a Bayesian estimation scheme dubbed Dynamic Causal Modeling (DCM). Following this work, Galka et al. (2008), Sauer et al. (2008), Daunizeau et al. (2009), and Freestone et al (2011) developed the estimation frameworks for continuum neural field equations that explain the richer dynamics of spatiotemporal neural fields.

>>>>

\section{Introduction}
\IEEEPARstart{T}{he} human brain has an inherent multiresolution architecture. Spatial scales of information processing and transfer range from individual proteins in synapses, to neurons, to networks of neurons. The multiresolution cortical architecture poses major modeling challenges to efficiently describe the brain's dynamics. This paper introduces a multiresolution data-driven framework for neural field modeling, the multiresolution approximate integrodifference equation (MRAIDE), to address this challenge.

Neural field models typically describe the meso and macroscopic dynamics of the human brain, but are also descriptive of finer-scale neurodynamics. The ability to create patient-specific neural models will contribute to our knowledge base of poorly understood diseases such as epilepsy, and furthermore, may also enable the development of new treatment strategies. This is particularly poignant with the advent of new devices that use therapeutic electrical stimulation. Presently, stimulation strategies for devices using corrective or modulating electrical stimulation operate in an open-loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists an enormous potential to improve the performance of these devices using systems theory. The design of a stimulation strategy using systems theory requires a model. Neural mass and neural field models are the ideal candidates for estimation and control algorithms. It is expected that parameters of the neural fields models will be patient specific and will therefore need to be inferred from data.

Creating patient specific neural models has been approached from a number of different standpoints. \dean{need to add proper citations here.} Valdes et al. (1999) used the neural mass model of Lopes Da Silva (1976) and Zetterberg et al. (1978) to fit EEG data. This approach was extended by David and Friston (2003) to coupled neural masses (using the equations of Jansen and Ritt (1995)) through a Bayesian estimation scheme dubbed Dynamic Causal Modeling (DCM). Following this work, Galka et al. (2008), Sauer et al. (2008), Daunizeau et al. (2009), and most recently Freestone et al (2011) developed the estimation frameworks for continuum neural field equations that explain the richer dynamics of spatiotemporal neural fields.
